Z-Resonance at LEP

### 2010-10-11 Z0-Publikation/tex/paper-template.tex
mein teil sozusagen fertig

For a detailed analysis of the mentioned processes it is important to be able
to distinguish between muonic and hadronic decays. By Monte Carlo studies the
differences between both processes could be figured out. This unlocks the
possibility to apply cuts on the measured distributions.

In this paper first results of measurements at center of mass energies of
\num{89.48}, \num{91.33} and \SI{93.02}{\giga\electronvolt} can be presented.

@@

Due to inefficiencies in energy measurement it is possible that not all the
energy provided by the apparatus can finally be seen. Therefore there are large
uncertainties in jet energy measurements, whereas these uncertainties are
smaller for muonic events. This fact can also be used to distinguish these
events from each other and from background as well. For example if the $Z^0$
decays into two very unstable tauons, these tauons decay very fast. Then two
neutrinos are created, but they cannot be measured, so that their contribution
to the visible energy is missing. The visible part of the tau decays resembles
hadronic events, so it is important to be able to separate this background from
the hadronic signal.

@@

\subsection{Hadronic events}
As it is now known how to extract hadronic events out of the data, it is
possible to determine the hadronic cross section using the following formula:
\begin{equation}
σ_{\mathrm{had}} = \frac{N}{Lε},
\end{equation}
where ε is the efficiency of the cuts. This efficiency could be calculated
using the following equation:
\begin{equation}
ε = \frac{N’_{\mathrm{MC}}}{N_{\mathrm{MC}}},
\end{equation}
where $N’_{\mathrm{MC}}$ is the number of Monte Carlo events that passed the
hadron cuts and $N_{\mathrm{MC}}$ is the total number of events in the Monte
Carlo file.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
energy $\sqrt{s}$ [\si{\GeV}] & hadronic events & cross section
$σ^{\mathrm{had}}$ [\si{\nano\barn}]\\
\hline
89.48 & 1848 ± 43 ± 52 & 10.55 ± 0.18 ± 0.16 \\
91.33 & 3980 ± 63 ± 79 & 29.98 ± 0.49 ± 0.43 \\
93.02 & 2149 ± 46 ± 51 & 14.56 ± 0.24 ± 0.24 \\
\hline
\end{tabular}
\vspace*{0.3cm}
\caption{\baselineskip=0.38cm analyzed hadronic events at different center of mass energies.}
\label{tab:hadr_events}
\end{center}
\vspace*{-0.5cm}
\end{table}

% These cross sections have then been used as data for a fit (see \fref{fit_hadrons}).
These cross sections have been used then to fit the Breit-Wigner distribution as
seen in \fref{fit_hadrons}.
By fitting the data
points the overall hadronic cross section $σ_0^{\mathrm{had}}$, the mass $M_Z$
and the decay width $Γ_Z$ of the $Z^0$-boson could be determined. The results
can be found in \tref{results}.

@@

On the other hand, muons could be distinguished from hadronic events as well as
from background. Therefore an analysis on muonic events is possible. This is
described in this section.

Firstly, the cross sections at the different center of mass energies could be
calculated in an analogue way to the hadronic events. The results of this can
be found in \tref{muon_events}.

@@

The achieved results by now offer furthermore an opportunity to obtain the
weak mixing angle using the formula
\begin{eqnarray}
Γ_l &=& \frac{G_F M_Z^3}{24 \sqrt{2}π}\left(1+[1-4|Q_l|\sin² Θ_W]²\right)\,.
\end{eqnarray}
As $Γ_l$ as well as $M_Z$ is known, the expression $\sin² Θ_W$ can be computed.
\begin{eqnarray}
\sin^2\Theta_W &=& \frac{1-\sqrt{\frac{\Gamma_e\cdot 24\sqrt{2}\pi}{G_FM_Z^3}-1}}{4}\\
\label{eqn:gammaquark}
\nonumber &=& 0.23 \pm 0.05 \pm 0.08
\end{eqnarray}
Thereby the weak mixing angle is determined, too.
\begin{eqnarray}
Θ_W &=& \arcsin(\sqrt{\sin^2\Theta_W})\\
\nonumber &=& 0.50 \pm 0.11 \pm 0.07\\
\nonumber &=& (28 \pm 6 \pm 4)\symbol{23}
\end{eqnarray}

@@

In reference to the standard model there is a way to compute $N_C$.
\begin{equation}
 Γ_{\mathrm{had}} = N_C \cdot K_{QCD}\cdot ( 2Γ_u + 3Γ_d )
 \label{eqn:gammahad2}
\end{equation}
The decay widths $Γ_u$ and $Γ_d$ can be calculated by the standard model, using the
value of the Weinberg angle as seen in \eref{gammaquark}. The charges of the up-quark was
set to $\frac{2}{3}$ and the charge of the down-quark to $-\frac{1}{3}$. Then the colour
factor can be calculated.
\begin{equation}
 N_C = 3.20 \pm 0.11 \pm 0.07
\end{equation}
This result confirms, that there are no new quarks to be expected with a mass
lower than some \SI{90}{\GeV}.

@@

\section{ Summary}
In order to sum up the results one has to say that this is a very precise
confirmation of the predictions of the standard model. The properties of the
predicted $Z^0$-boson could be measured precisely. Additionally the first exact
measurement of the weak mixing angle was done successfully, which was one of the
most important tasks of this experiment.

>>>>

For a detailed analysis of the mentioned processes it is important to be able
to distinguish between muonic and hadronic decays. By Monte Carlo studies the
differences between both processes could be figured out using pure hadronic and
muonic data samples as well as simulated data.
% , which was necessary
% because the hadronic and muonic cross sections and decay widths are going to
% be calculated seperately.
This facilitates the
possibility to apply cuts on the measured distributions.

In this paper first results of measurements at center of mass energies of
\num{89.48}, \num{91.33} and \SI{93.02}{\giga\electronvolt} will be presented.

@@

Due to inefficiencies in energy measurement it is possible that not all the
energy provided by the apparatus can finally be seen. Therefore there are large
uncertainties in jet energy measurements, whereas these uncertainties are
smaller for muonic events. Nevertheless also for these events there is a
statistical uncertainty limiting the energetic resolution. This fact can also be
used to distinguish these
events from each other and from background as well. For example if the $Z^0$
decays into two very unstable tauons, these tauons decay very fast. Then two
neutrinos are created, but they cannot be measured, so that their contribution
to the visible energy is missing. The visible part of the tau decays resembles
hadronic events, so it is important to be able to separate this background from
the hadronic signal.

@@

\subsection{Hadronic events}
As it is now known how to extract hadronic events out of the data, it is
possible to determine the hadronic cross section using the following formula:
\begin{equation}
σ_{\mathrm{had}} = \frac{N}{Lε},
\label{eqn:sigma}
\end{equation}
where ε is the efficiency of the cuts. This efficiency could be calculated
using the following equation:
\begin{equation}
ε = \frac{N’_{\mathrm{MC}}}{N_{\mathrm{MC}}},
\end{equation}
where $N’_{\mathrm{MC}}$ is the number of Monte Carlo events that passed the
hadron cuts and $N_{\mathrm{MC}}$ is the total number of events in the Monte
Carlo file.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline
energy $\sqrt{s}$ [\si{\GeV}] & hadronic events & cross section
$σ^{\mathrm{had}}$ [\si{\nano\barn}]\\
\hline
89.48 & 1848 ± 43 ± 52 & 10.55 ± 0.18 ± 0.16 \\
91.33 & 3980 ± 63 ± 79 & 29.98 ± 0.49 ± 0.43 \\
93.02 & 2149 ± 46 ± 51 & 14.56 ± 0.24 ± 0.24 \\
\hline
\end{tabular}
\vspace*{0.3cm}
\caption{\baselineskip=0.38cm analyzed hadronic events at different center of
mass energies. The first given error is the statistical, the second one is the
systematic uncertainty.}
\label{tab:hadr_events}
\end{center}
\vspace*{-0.5cm}
\end{table}

The systematic uncertainties listed in this table are derived from the
systematic uncertainties of $ε_{\mathrm{had}}$ and the influence of variations
of the cuts.

% These cross sections have then been used as data for a fit (see \fref{fit_hadrons}).
The listed cross sections have been used then to fit the Breit-Wigner
distribution as
seen in \fref{fit_hadrons}.
By fitting the data
points the overall hadronic cross section $σ_0^{\mathrm{had}}$, the mass $M_Z$
and the decay width $Γ_Z$ of the $Z^0$-boson could be determined. The results
can be found in \tref{results}.

@@

Muonic signals could be distinguished from hadronic events as well as
from background. Therefore an analysis on muonic events is possible. This is
described in this section.

Firstly, the cross sections at the different center of mass energies could be
calculated in an analogue way to the hadronic events (see \eref{sigma}). The
results of this can
be found in \tref{muon_events}.

@@

The achieved results by now offer furthermore an opportunity to obtain the
weak mixing angle using the formula
\begin{eqnarray}
Γ_l &=& \frac{G_F M_Z^3}{24 \sqrt{2}π}\left(1+[1-4|Q_l|\sin² Θ_W]²\right)\,.
\end{eqnarray}
As $Γ_l$ as well as $M_Z$ is known, the expression $\sin² Θ_W$ can be computed.
\begin{eqnarray}
\sin^2\Theta_W &=& \frac{1-\sqrt{\frac{\Gamma_e\cdot 24\sqrt{2}\pi}{G_FM_Z^3}-1}}{4}\\
\label{eqn:gammaquark}
\nonumber &=& 0.23 \pm 0.05 \pm 0.08
\end{eqnarray}
Thereby the weak mixing angle is determined, too.
\begin{eqnarray}
Θ_W &=& \arcsin(\sqrt{\sin^2\Theta_W})\\
\nonumber &=& 0.50 \pm 0.11 \pm 0.07\\
\nonumber &=& (28 \pm 6 \pm 4)\symbol{23}
\end{eqnarray}
This angle is, as a parameter of the electroweak theory, a fundamental
constant describing this interaction as a symmetry group.

@@

In reference to the standard model there is a way to compute $N_C$.
\begin{equation}
 Γ_{\mathrm{had}} = N_C \cdot K_{QCD}\cdot ( 2Γ_u + 3Γ_d )
 \label{eqn:gammahad2}
\end{equation}
The decay widths $Γ_u$ and $Γ_d$ can be calculated by the standard model, using the
value of the Weinberg angle as seen in \eref{gammaquark}. The charges of the up-quark was
set to $\frac{2}{3}$ and the charge of the down-quark to $-\frac{1}{3}$. Then the colour
factor can be calculated.
\begin{equation}
 N_C = 3.20 \pm 0.11 \pm 0.07
\end{equation}
This result confirms, that there are no new quarks to be expected with a mass
lower than some \SI{45}{\GeV}.

@@

\section{ Summary}
In order to sum up the results one has to say that this is a first
confirmation of the predictions of the standard model. The properties of the
predicted $Z^0$-boson could be measured precisely. Additionally the first exact
measurement of the weak mixing angle was done successfully, which was one of the
most important tasks of this experiment.
